---
title: "hw3"
output: html_document
date: '2023-10-10'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

This data set consists of 1,384,617 observations of 15 variables. Some key variables include product ID, department, and product name. The user ID's appear to be grouped by order number, which makes sense because when you order on Instacart, you are given a unique order number that corresponds to your unique user ID. There are 134 aisles, with fresh vegetables and fresh fruits being the most ordered items s
```{r}
library(tidyverse)
library(dplyr)
library(p8105.datasets)
data("instacart")

instacart = 
  instacart |> 
  as_tibble()
  
instacart |> 
  count(aisle) |> 
  arrange(desc(n))  

instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()

instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```

## Problem 2

```{r}
data("brfss_smart2010")

## data cleaning 

  brfss_smart2010 =
  brfss_smart2010 |>     
  janitor::clean_names() |> 
  filter(
    topic == "Overall Health",
    response %in% c("Excellent","Very good", "Good", "Fair", "Poor")) |> 
  mutate( 
    response = 
      fct_relevel(response, c("Poor", "Fair", "Good","Very good","Excellent" ))
  )
  brfss_smart2010


```

## in 2002, which states were observed at 7 or more locations? What about in 2010?
In 2002, 6 states were observed at 7 or more locations. In 2010, 14 states were observed in 7 or more locations. 

```{r}
states_obs <- brfss_smart2010 |> 
  filter(year %in% c(2002, 2010)) |> 
  select(year, locationabbr, locationdesc) |> 
  distinct() |> 
   group_by(year, locationabbr) |>
  summarize(n = n()) |> 
  filter(n >= 7) |> 
  arrange(year, desc(n))
```

## Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state.

```{r}  
excellent_data <- brfss_smart2010 |> 
  filter(response %in% c("Excellent")) |>
  select(year, locationabbr, data_value) |>
  group_by(locationabbr, year) |> 
  mutate(
    avg_data_value = mean(data_value)) |> 
  select(year, locationabbr, avg_data_value) |> 
  distinct()
```

## spaghetti plot

The plot shows the average data value for each state from the years 2002-2010. Overall, the data has a lot of overlap, but has some outlier values. For example, there appears to be a low average data value score in 2005 for the state of West Virginia. 
```{r}
excellent_data |> 
  ggplot(aes(x = year, y = avg_data_value)) +
  geom_line(aes(group = locationabbr, color = locationabbr)) +
  labs(
    title = "Average Data Value Among States by Year",
    x = "Year",
    y = "Average Data Value")

```

## make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State

Overall, there is more variability in the 2010 resonses compared to the 2006 responses. The "very good" response appears to have improved the most over the course of four years. 
```{r}
brfss_smart2010 |> 
  select(year, data_value, locationabbr, locationdesc, response) |> 
  filter(
    year == 2006 | year == 2010,
    locationabbr == "NY") |> 
  ggplot(aes(x = data_value, fill = response)) +
  geom_boxplot(aes(fill = response), alpha = .5) +
   
  labs(
    title = "Data Value Across Responses in New York",
    x = "Data Value",
    fill = "Response") + 
  facet_grid(. ~ year)
```

## Problem 3 

```{r}
## load, tidy, merge, and otherwise organize the data sets.

covar_data <- read.csv("nhanes_covar.csv", skip = 4)  #, sep = '\t', skip = 4, col.names = c("seqn", "SEQN"))
accel_data <- read.csv("nhanes_accel.csv")

covar_data_clean = covar_data |> 
  filter(age >= 21) |> 
  drop_na() |> 
  mutate(
    sex = 
      case_match(
        sex,
        1 ~ "male",
        2 ~ "female"
      ), 
    sex = as.factor(sex),
    education = 
      case_match(
        education,
        1 ~ "less than high school",
        2 ~ "high school equivalent",
        3 ~ "more than high school"
      ), 
    education = as.factor(education)
  )

merged_data = 
  left_join(covar_data_clean, accel_data, by = "SEQN")

```

## Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items.

Between 20-40 years old, most individuals had more than a high school education. Between 40-60 years old, the education level is more variable between high school equivalent, less than high school, and more than high school. Ages 60 and above, less individuals have more than a high school education compared to the other age groups. 
```{r}
merged_data |>
  drop_na() |> 
  group_by(sex, education) |>
  summarize(count = n()) |> 
  knitr::kable(digits = 1)  

merged_data |>
  drop_na() |> 
  ggplot(aes(x = age, fill = education)) +
  labs(
    title = "Age Distributions Across Education Levels") +
  geom_histogram()
```
## Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities 

Overall across the plots, it does not seem that education level impacts total activity. The more than high school panel has the most data points, but they are sporadic like the other two education level panels. Across sex, women appear to have slightly more total activity than men in the high school equivalent and more than high school panels. 
```{r}
merged_data |> 
  mutate(
    total_activity = rowSums(across(min1:min1440), na.rm = TRUE)) |> 
    ggplot(aes(x = age, y = total_activity, color = sex)) + geom_point() + geom_smooth(se = FALSE) + labs(y = "Total Activity") + facet_grid(.~education)
  
```

## Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex.
Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.

The graph illustrates that across all education levels, females had a higher average 24-hour activity rate compared to males. Females with an education level of more than high school have the highest median 24-hour activity compared to females in the other education levels. Males with an education level of less than high school have the highest median 24-hour actvivity compared to males in the other education levels. 
```{r}
day_activity = merged_data |> 
  group_by(education, sex) |> 
  summarize(across(starts_with("min"), ~ mean(.), .names = "mean_{.col}"), .groups = "drop") |> 
  pivot_longer(cols = starts_with("mean_"), names_to = "time", values_to = "mean") |> 
  mutate(
    time = substring(time, 9), 
    time = as.numeric(time))
ggplot(day_activity, aes(x = time, y = mean, fill = sex)) + geom_boxplot() +  facet_grid(.~ education) +
  labs(title = "24-Hour Activity Time Courses by Education Level and Sex")
```

